---
title: "Report"
author: "Yunran Chen"
date: "2017/12/10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1.Summary

## 2.Introduction

Uterotropic bioassay is one approach for screening chemicals for endocrine disrupting effect. An international mutilaboratory study was conducted with this method using known estrogen agonist(`EE`) and antagonist(`ZM`). The main goal is to assess whether the results were consistent across the laboratories. For one aspect, we care about:1)whether the method is successful for identifying endocrine disrupting effect of `EE` and `ZM`; 2)how dose response? is there a "change" dose level of `EE` for endocrine disrupting effect; 3)do the protocols differ in their sensitivity to detect estrogenic and anti-estrogenic effects; which one is recommended? For the other aspect, we care about:1) whether the results for abovementioned problems vary across labs;2)which labs stand out for being different?

The dataset including: response variable `uterus`(Uterus weight (mg)); key explanatory variables: `EE`(Dose of estrogen agonist,mg/kg/day),`ZM`(Dose of estrogen antagonist, mg/kg/day);other explanatory variables:  `Lab`(19 Laboratory at which assay was conducted), `Protocol`(4 levels factor, representing immature female rats dosed by oral gavage (3 days), immature female rats dosed by injection (3 days), adult ovariectomized female rats dosed by injection (3 days), adult ovariectomized female rats dosed by injection (7 days) seperately.); `Weight`(Body weight of rat (g)),`Group` (Lab replicate group (6 rats were used per group)).

The characteristics of the dataset.

## 3.Methods and Model

### 3.1 EDA

From the plot, we get three main information on the dataset.1)linear regession assumption.(quadratic form)
2)There are significant interaction among variables. Because each labs adopted differet treatment. So we have to include the iteraction into our regression model.
3)Many discrete variables. We need to keep the model simple.

I chose boxplot to illustrate the relationship between continuous variable (`uterus`) and discrete variable(`EE` and `ZM`).Different colors represent different `labs`, and I use facet to represent different `protocol`. From the plot1, we can see that: 1)The trend for expection looks like quadratic form; But overall, the uterus decrease as the zm increase; The variance under different level of `ZM` is different. The variance of `ZM`=0 is much larger than the other two groups; 2)Overall, the boxplots look similar. However, the different labs act different especially when `ZM`=0. Some labs, like TNO in `protocol="B"`,`ZM=0.1` acts like an "outlier". There is interaction between `lab` and `ZM`. 3)Each labs adopted different treatment. So for some case, limited labs conducted the study. Similar for the case `EE`.

I chose scatterplots with regression line to illustrate the relationship between two continuous variable (`uterus` and `weight`). Because the significant differece on `weight` between immature rats and adult rats, I use facet to represent different `protocol`. Notice the range for y-axis is different. We can see that: 1)For immature rats, `weight` seems do not effect the `uterus`. While for adult rats, the `uterus` decrease as the `weight` increase. There is significant interaction.

### 3.2 Model

I build a linear regression model with all variables except `group`. Apart from `weight` and `uterus`, I consider all other variables as factor. Also, I include the interaction between estrogen chemicals and different labs along with protocols. Notice for the factors need to use a set of dummy variables to represent, so there are 218 predictors including the intercept.

$Y=intercept+\beta_1EE+\beta_2ZM+\beta_3lab+\beta_4protocol+\beta_5weight+\beta_6EE:protocol+\beta_7ZM:protocol+\beta_8EE:lab+\beta_9ZM:lab+\epsilon$ 

where $\epsilon \sim N(0,\sigma^2)$

Notes:

1. The value of `EE` and `ZM` is discrete. It is resonable to set it as factor. And compared to the model with numeric `EE` and `ZM`, the model I chose performed better in fitting the model(Ajusted R-square 0.9538; the model include numeric `EE` and `ZM` and quadradic form can only achieve 0.8286) and get the results as I expected from the boxplot.(`ZM:labs` is significant in F-test)

2. The model does not include `group` for it does not pass the F-test. This make sense, which also prove that the study is replicable.

3. Notice that each labs adopted different schemes for the study.(See the jitter plot) So interaction would be significant among all the variables (except for `group`). However, there are 19 labs and 8 different levels for `EE`. We need to be careful when include the interaction. I adopted the rule "keep the simplest model" and make sure that each variable and interaction I include in is super significant.(with p-value for F-test less than 2.2e-16).

### 3.3 Methods

I use MLE/OLS to estimate the parameters. I use F-test, t-test for a single coefficient and t-test for linear combination of coefficients to answer the ques we care about. Actually, all the questions can be answer by only two plots.(See the Page 3 for details.)

a.1 Is the uterotrophic bioassay successful overal at identifying effects of EE and ZM

Using F-test. 

And check all the sign for the coefficients `EE` (positive) and `ZM` (negative). 

Hypothesis: $H_0: \beta_{11}=\beta_{12}=\beta_{13}=0$(failed to detect) v.s. $H_1:\exists \beta_{1i} \neq 0$(detect successfully)

statistics: F-stat(using anova() in R)

$F=\frac{\parallel (P_k - P_{k-1})Y \parallel ^2/(r(P_k)-r(P_{k-1}))}{\hat{\sigma}^2} \sim F(r(P_k)-r(P_{k-1}),n-p)$

Note: here the $\beta{1i}$ means all the coefficients before `EE`. Similar F-test for `ZM`

a.2 Do some labs fail to detect such effects?

Constructing t-test for linear combination of coefficients.Take lab Bayer for example. The sum of coefficients before `EE` (`EE` here including `EE0.01`, `EE0.03`,...) and `EE:labBayer` is t-distrbution. All also, we need to check the sign for significant coefficients.

Hypothesis: $H_0: \Sigma_i (\beta_{1i}+\beta_{1i:labBayer})=0$(lab Bayer failed to detect) v.s. $H_1:\Sigma_i (\beta_{1i}+\beta_{1i:labBayer}) \neq 0$(lab Bayer detect successfully)

statistics: t-stat(need construct function in R)

$t=\frac{\lambda^T\hat{\beta}-\lambda^Tb_0}{\hat{\sigma}\sqrt{\lambda^T(X^TX)^{-1}\lambda}} \sim t(n-p,0,1)$

where $\lambda_i=1$ for the position `EE` or `EE:labBayer`,else $\lambda_i=0$.

a.3 The change level of dose for the effect? Does this vary across labs?

Using t-test for single coefficient and find the change point from insignificant to significant. For the first question, we concerntrate on coefficients of `EE`, `ZM`. For the second question, for each $lab_i$ we concerntrate on coefficients of `EE:lab_i`, `ZM:lab_i`.

Hypothesis: $H_0: \beta_{1i}=0$ v.s. $H_1:\beta_{1i} \neq 0$

statistics: t-stat

$t=\frac{\hat{\beta}-b_0}{\hat{\sigma}\sqrt{(X^TX)^{-1}}} \sim t(n-p,0,1)$

b Does the dose response vary across labs? If so, which one stands out as being different?

Using t-test for single coefficient.For each $lab_i$ we concerntrate on coefficients of `EE:lab_i`, `ZM:lab_i`. For certain $dose_j$, if some of them are significant, some are not. We would say the dose response vary across labs. If some of the pvalue are especially small, those labs may stand out as being different.

c Do the protocols differ in their sensitivity to detect the effects? Is there one can be recommend?

Using F-test for the interaction `EE:protocol`, `ZM:Protocol`.If they are significant, we may say the protocols differ. 

A good protocol should have following properties: 1)Robust. The variance of coefficients of `EE:protocol_i` and `ZM:protocol_i` should be small. 2)Significant.The absolute value of coefficients before `EE:protocol_i` and `ZM:protocol_i` should be large, so that we can easy to detect. 3)Consistent. Under the protocol, different labs should get similar result.  The 1)3) can be settled down by F-test. 2) can be settled down by comparing the estimator.(the location of confidence interval)

Notes:

1. The main problem of this dataset is too much factors and interactions leading to too much predictor. If adopted Bayesian methods, it would be time consuming.(There are some improvements--adding selection process in Appendix) So I adopted MLE/OLS for it is simple and fast.

2. For overall test for discrete variable (factors), I adopted F-test. And for specific problem, I use t-test.

### 3.4 Outliers and Influential points

According the residuals, there are three possible "outliers"(1426,926,1586). However, I text these outliers in the boxplots. I find that points 1586,926 are outliers with $ZM=0,EE=0.1$. In the boxplot, they are much higher compared to all other points. So I consider it as an outlier. However, the points 1426 is with $ZM=1,EE=3$. It maybe the result of the interaction of `EE` and `ZM`.And not much labs conduct studies under this treatment. Only a few studies including both `EE` and `ZM`. So I keep the points 1426.

Notices: all the outliers appears in protocol C,D. This suggest us these procotols are not recommended.

## 4.Results

```{r library, message=FALSE, warning=FALSE,echo=FALSE}
library(tibble)
library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(stringr)

bioassay.read = read.table("http://stat.duke.edu/sites/stat.duke.edu/files/bioassay.txt",
                       header=T,stringsAsFactors = FALSE) %>% as.tibble()

bioassay=bind_cols(map_df(bioassay.read %>% select(uterus,weight,EE,ZM),~.x %>% as.numeric(.)),
                   map_df(bioassay.read %>% select(protocol,lab,group),~.x %>% as.factor(.))
                   ) %>% as.tibble()
bioassay.fac=bioassay %>% mutate(EE=as.factor(EE),ZM=as.factor(ZM))
```

```{r summary}
##1)We can consider `weight` and `uterus` as continuous variable. All other variables only have seperate values.
summary(bioassay)
bioassay %>% select(EE) %>% table()
bioassay %>% select(ZM) %>% table()
##2)There are significant interaction among variables. Because each labs adopted differet treatment.
table(bioassay$EE,bioassay.fac$lab)
##visualization for points
#ggplot(data=bioassay,mapping = aes(y = lab,x = protocol,color=as.factor(ZM)))+
#  geom_jitter(alpha=0.5)
#ggplot(data=bioassay.fac,mapping = aes(y = lab,x = ZM,color=protocol))+
#  geom_jitter(alpha=0.5)
#ggplot(data=bioassay,mapping = aes(y = lab,x = protocol,color=as.factor(EE)))+
#  geom_jitter(alpha=0.5)
#ggplot(data=bioassay,mapping = aes(y = lab,x = protocol,color=as.factor(EE)))+
#     geom_count(alpha=0.5,position = "jitter")
```


```{r EDA}
##---------------------boxplot
##different labs have diff mean on ZM. interaction on labs and protocol.
##when consider ZM as numeric, the diff is not sig in pic. but in factor - diff.
###4var:ZM

outliers=c(1426,926,1586)#,1425,1946,2666)
ggplot(data=bioassay.fac,mapping = aes(y = uterus,x = ZM,color=lab))+
  geom_boxplot()+theme_bw()+facet_wrap(~ protocol)+
  geom_text(data=bioassay.fac[outliers,],
            aes(label=outliers,color=lab))
##caculate the mean for different group 
#bioassay.fac %>% group_by(ZM) %>% summarize(mean=mean(uterus,na.rm=T))

##4var:EE
ggplot(data=bioassay.fac,mapping = aes(y = uterus,x = EE,color=lab))+
  geom_boxplot()+theme_bw()+facet_wrap(~ protocol)+
  geom_text(data=bioassay.fac[outliers,],
            aes(label=outliers,color=lab))
##caculate the mean for different group.
#bioassay.fac %>% group_by(EE) %>% summarize(mean=mean(uterus,na.rm=T))

##weights
ggplot(data=bioassay.fac,mapping = aes(y = uterus,x = weight))+
  geom_point()+geom_smooth(method = "lm")+theme_bw()+facet_wrap(~ protocol,scales = "free")+
  geom_text(data=bioassay.fac[outliers,],
            aes(label=outliers,color=lab))

```


```{r lm.full.fac}
lm.full.fac=lm(data = bioassay.fac[-c(1586,926),],formula = uterus~EE+ZM+lab+protocol+weight+EE:protocol+ZM:protocol+EE:lab+ZM:lab) 

plot(lm.full.fac) ##use the plot to check model assumption. (detect outliers, normality, influential points.)

#bioassay.fac %>% select(EE,ZM) %>% table() -- there is imbalanced distribution 

summary(lm.full.fac)
#Adjusted R-squared:  0.9538 

##if consider EE,ZM as numeric
#lm.full.q=lm(data = bioassay,formula = #uterus~poly(EE,2)+poly(ZM,2)+lab+protocol+weight+EE:protocol+ZM:protocol+EE:lab+ZM:lab)
#anova(lm.full.q)
#summary(lm.full.q) #0.8286 #p=69 
```

```{r functions}
eff.tbl=function(lm.obj){
  
  ind.mat=matrix(0,nrow=nrow(summary(lm.obj)$coefficients),ncol=7)
  colnames(ind.mat)=c("EE","ZM","lab","protocol","EEdose","ZMdose","interaction")
  ind.mat=cbind(coef=summary(lm.obj)$coefficients%>%rownames(.),ind.mat)
  ind.mat[str_detect(ind.mat[,1],"EE"),"EE"]=1
  ind.mat[str_detect(ind.mat[,1],"ZM"),"ZM"]=1
  ind.mat[str_detect(ind.mat[,1],"lab"),"interaction"]=1
  ind.mat[str_detect(ind.mat[,1],"protocol"),"interaction"]=1
  
  for (dose in c(bioassay.fac$EE %>% levels())){
     ind.mat[str_detect(ind.mat[,1],paste0("EE",dose)),"EEdose"]=dose
  }
  for (dose in c(bioassay.fac$ZM %>% levels())){
     ind.mat[str_detect(ind.mat[,1],paste0("ZM",dose)),"ZMdose"]=dose
  }
  for (lab in c(bioassay.fac$lab %>% levels())){
     ind.mat[str_detect(ind.mat[,1],paste0("lab",lab)),"lab"]=lab
  }     
  for (protocol in c(bioassay.fac$protocol %>% levels())){
     ind.mat[str_detect(ind.mat[,1],paste0("protocol",protocol)),"protocol"]=protocol
  }  
  ind.tbl=ind.mat%>% as.tibble()
  return(ind.tbl)
}

t.test=function(lm.obj,str.ee,str.lab,str.ori){
ind.tbl=eff.tbl(lm.obj)
cov.coef=vcov(lm.obj)
p=nrow(ind.tbl)
i.levels=bioassay.fac %>% pull(str.ori) %>% levels() #original colnames in bioassay.fac
i.n=i.levels %>% length(.)
lambda=matrix(nrow=i.n,ncol=nrow(ind.tbl))
lab.ee=t.value=p.value=numeric(i.n)

for (i in 1:i.n){
  lambda[i,]=((ind.tbl[str.lab]==i.levels[i])&(ind.tbl[str.ee]=="1"))|((ind.tbl[str.ee]=="1")&(ind.tbl["interaction"]=="0"))
  lab.ee[i]=summary(lm.obj)$coefficients[lambda[i,],"Estimate"] %>% sum()
  denominatior=(cov.coef[lambda[i,],lambda[i,]]) %>% sum() %>% sqrt()
  nominatior=lab.ee[i]
  t.value[i]=abs(nominatior/denominatior)
  p.value[i]=pt(q = t.value[i],df = summary(lm.obj)$df[2],lower.tail = FALSE)
}

res=tibble(i.levels,p.value,estimator=lab.ee)
return(res)
}
```



```{r result}
#0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

##a.1
anova(lm.full.fac)

coefs=rownames(summary(lm.full.fac)$coefficients)

res=summary(lm.full.fac)$coefficients %>%
  cbind(coef=coefs,.) %>% 
  as.tibble() #get the coefficients matrix

colnames(res)=c("coef","estimate","std_error","t_value","p_value")

res %>% filter(as.numeric(p_value)<0.05) %>% slice(1:10)
##uterotrophic bioassay successful overal at identifying effects of EE and ZM. F-test for EE,ZM are significant. For the significant coefficients, all EE are positive, ZM are negative. 

##a.2 do some labs fail t detect such effects
##There are some labs fail to detect such effects, as follows:"Huntingd""Bayer""ChungKor""TNO""Zeneca"  
a.2.ee=t.test(lm.full.fac,"EE","lab","lab") 
a.2.ee
a.2.ee %>% filter(((p.value<0.05)&(estimator<0))|(p.value>0.05))# %>% pull(i.levels)

a.2.zm=t.test(lm.full.fac,"ZM","lab","lab")
a.2.zm
a.2.zm %>% filter(((p.value<0.05)&(estimator>0))|(p.value>0.05))#%>% pull(i.levels)

##a.3 the change dose for ee? vary across labs?
##From the output 
```

## 5.Conclusions

## Part I: OLS/MLE


```{r lm.full.2}

###------Q1
#ind=str_detect(summary(lm.full.simple)$coefficients %>% rownames(),"EE")
#eff.coeff.2=summary(lm.full.simple)$coefficients %>%
#  as.data.frame() %>% 
#  mutate(coeff=rownames(.)) %>% 
#  filter(ind)

```



```{r lm.full.fac.1}




labeff("EE","lab","lab")
labeff("ZM","lab","lab")
#(3)dose
labeff("EE","EEdose","EE") #0.1-|-0.3
labeff("ZM","ZMdose","ZM") #0-|-0.1
#b.(1)dose+lab

#b.(2)

#c.(1)protocol
labeff("EE","protocol","protocol")
labeff("ZM","protocol","protocol")
```

